---
title: "Carrefour Kenya Marketing Strategies Part 4"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Business Understanding

As a Data analyst at Carrefour Kenya, I undertook a Data Science Project to inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax). My project has been divided into four parts where I'll explore a recent marketing dataset by performing various unsupervised learning techniques and later providing recommendations based on my insights.

## Data Understanding

```{r}
# Loading the necessary libraries and automatically installing them if not present

pacman :: p_load(rio,tidyverse, Amelia ,anomalize)
```

```{r}
# Previewing our dataset

data<-import("~/R/Unsupervised Learning Week 14 IP/Supermarket_Sales_Forecasting - Sales.csv")
data
```


``` {r}
# Checking the size and shape of data
dim(data)
```

``` {r}
# Viewing data types using str().

str(data)
```

## Data Preparation
``` {r}
# Checking for missing data in our data
missmap(data)
```
``` {r}
# Checking for duplicated records

duplicated_rows <- data[duplicated(data),]  
duplicated_rows # No duplicates in our data
```
``` {r}
# Converting all column names to lowercase and date feature to datetime

# changing case
names(data)[1:2] <- tolower(names(data)[1:2])
```

```{r}
# converting to datetime
data$date <- as.POSIXct(data$date, "%m-%d-%Y",tz = "GMT")
```

``` {r}
# Checking for Outliers in Sales

boxplot(data[,c(2)]) # Summary statistics will tell us if these few points are outliers
```
## Exploratory Data Analysis (Univariate & Bivariate )

# Univariate

```{r}
summary(data)

# Since we want to detect anomalies or fraud in sales data, we'll let these points remain in the data which reveals maximum sale of 1042.65 and minimum of 10.68
```

```{r}
# Histogram of sales 

# Distribution of Sales
hist(data$sales,
        main = "Distribution of Sales(Benford's Law)",
        xlab = "Sales",
        ylab = "Frequency")
```

```{r}
# Similar to part 1 of this project, the histogram seems to be somewhat similar to Benford's Law Distribution of anomalous numbers in real life numerical data. However it has some slight differences in the first bar which should be higher than the second one.
```

# Bivariate
```{r}
# First feature extracting the day and month from date and converting day to numeric

data$day <- format(data$date, "%d")
data$month <- format(data$date, "%m")
data$day <- as.numeric(data$day)

# Plotting scatter plot of date vs sales
plot(data[,c(1,2)])   # There seems to be no clear relationship between date and sales
```

```{r}
# Plotting scatter plot of month vs sales
plot(data[,c(4,2)])   # There seems to be no clear relationship between month and sales as well
```

```{r}
# Correlations

correlations <- cor(data[,2:3], method = "pearson")
round(correlations, 2)  # Confirms no relationship between day and sales
```

## Anomaly Detection
```{r}
# Groupind data by month
monthly_data <- data %>% 
          group_by(date,month) %>% summarise_at(vars(sales),funs(sum(.,na.rm=TRUE)))

# converting monthly_data to tibble
monthly_data <- as_tibble(monthly_data)

# Detecting Anomalies in monthly sales
monthly_data %>%
    time_decompose(sales) %>%
    anomalize(remainder) %>%
    time_recompose() %>%
    plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)

```
```{r}
# converting overall data to tibble
data_tibble <- as_tibble(data)

# Detecting Anomalies in overall sales
data_tibble %>%
    time_decompose(sales) %>%
    anomalize(remainder) %>%
    time_recompose() %>%
    plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)

```
```{r}
# From the two graphs above,it is clear that Carrefour Supermarket has no cases of fraud/anomalies in the sales data. Everything is okay. Cases of anomalies/fraud would have been highlighted by red dots

# Therefore it is true that its distribution of sales observes Benford's Law as seen in Univariate Analysis.
```


## Insights

```{r}
# Carrefour should continue with its protocols in recording sales data as it is clean and free of fraudulent activities.

# The sales of products is independent of the day or month. However, from part one of this analysis, the sales is dependent on the time of day. See first part of the project.

# We need sales data from each branch to conduct anomaly detection from each branch.
```